---
title: "Marine data prep"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(dplyr)
library(glue)
here::i_am("reports/00data_prep.Rmd")
ships <- readr::read_csv(here::here("data","ships.csv"))
```


## Objective of data prep

There is half a GB of information in `ships.csv` with multiple columns. The reasons for this data prep are:

1. combine the `SHIP_ID` column and the `SHIPNAME` column to add in the ship selector dropdown the possibilty to type the ship you are looking for with either the ID or the shipname.

2. make the file smaller by only grabbin the data that it is going to be in the app. The variables are: 

  - [x] `LAT` !!!
  - [x] `LON` !!!
  - [x] `SPEED`
  - [x] `DESTINATION`: is name of place
  - [x] `FLAG`
  - [x] `SHIPNAME` !!!
  - [x] `SHIP_ID`  !!!
  - [x] `SHIPTYPE` !!! .... is `ship_type` == `SHIPTYPE`? explore...
  - [x] `DATETIME`: minutes are very important
  - [x] `week_nb`
  - [x] `port`: is name of port. `PORT` is the same as `port` but the ladder is better suitable for data presentation
  - [x] `is_parked`

3. Finally do a quick load check between using `data.table` and `arrow`. For these will help the app load and do calcultions faster.

!!!: are essential for performing the basic tasks in the home assignment.

## Select columns

```{r selectCols}

key_vars_ship <- ships %>% 
  select(
    SHIP_ID,
    SHIPNAME,
    SHIPTYPE,
    ship_type,
    FLAG,
    DESTINATION, 
    port,
    SPEED,
    is_parked,
    DATETIME,
    week_nb,
    LAT,
    LON
  )
  
  
```



## About combining columns

After checking `ship_type` and `SHIPTYPE`, one is a character variable whereas the second is numerical variable. So i am going to combine these two too.

### combine ship type columns and ship name columns

```{r combineShiptypesAndnames}

clean_ship_selectors <- key_vars_ship %>%
  mutate(
    clean_ship_types = paste(ship_type, "-", SHIPTYPE),
    clean_ship_names = paste(SHIPNAME, "-", SHIP_ID)
  ) %>%
  select(
    clean_ship_names,
    clean_ship_types,
    FLAG,
    DESTINATION,
    port,
    is_parked,
    DATETIME,
    week_nb,
    LAT,
    LON,
    !c(ship_type, SHIPTYPE, SHIPNAME, SHIP_ID)
  ) %>%
  relocate(clean_ship_names, clean_ship_types, .before = FLAG)

```





## Clean column names

```{r clean_columns}

clean_ship_data <- clean_ship_selectors %>% 
  rename(
    "ShipName" = clean_ship_names,
    "ShipType" =  clean_ship_types,
    "week_of_year" = week_nb
  )
  
  
  
```


## Write new csv

```{r writecsv}

data.table::fwrite(clean_ship_data,
                   here::here("data", 
                              "clean_ships.csv"),
                   bom = TRUE)
```


## Test load times

```{r loadtime}
library(data.table)
library(arrow)

tictoc::tic(msg = "load with data.table")

datatable_ships <- fread(here::here("data", "clean_ships.csv"),
                         encoding = "UTF-8")
tictoc::toc()


tictoc::tic(msg = "load with arrow")

arrow_ships <- read_csv_arrow(here::here("data", "clean_ships.csv"))
tictoc::toc()

```



## Conclusion

9 columns were removed and columns were resorted, transformed, and renamed.

The clean dataframe has `r ncol(clean_ship_selectors)` columns and the same number of data points: `r nrow(clean_ship_selectors)`.

The data prep cut down the size of the csv from ~398.1MB to ~309.4MB. Data was reduced to `r 309.* 100 / 398.1` % of the original size.

For this demo, I am going to suppose I must use a csv file and cannot use any other file type. Thus, `data.table` will allow a faster load time
