---
title: "01Protyping_functions"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(data.table)
library(dtplyr)
library(dplyr, warn.conflicts = FALSE)
library(tictoc)

here::i_am("reports/01funct_prototyping.Rmd")
```

## Load data

```{r loadData}
tic(msg = "load and guess")
df_ships <- fread(
  here::here("data", "clean_ships.csv"),
                         encoding = "UTF-8"
)
toc()

tic(msg = "load and specify")

df_ships <- fread(
  here::here("data", "clean_ships.csv"),
  encoding = "UTF-8",
  sep = ",",
  header = TRUE
)

toc()

#dont make a copy of the df when calling mutate
lazy_df_ships <- lazy_dt(df_ships, immutable = FALSE)
```

Better to specify than not specify

## Make test data

```{r testData}
sample_ship_t_n <-
  lazy_df_ships %>% 
  select(ShipName, ShipType) %>% 
  distinct() %>%
  as_tibble() %>% 
  sample_frac(0.25) 
```


## Unique vessel types and names

Considering there are only 9 vessel types, this information can be extracted and created on the spot instead of querying distinct() from the loaded df. 

I will also extract unique vessel names to [**mock a coocked answer for the app**](#Mock answer), so the only loading time of the app would be over the loading of the big csv and shiny dependencies and not burden with these calculations.

```{r uniqueVessel}

Vessel_types_n_names <- lazy_df_ships %>%
  select(ShipType,ShipName) %>% 
  distinct() %>%
  as_tibble()

fwrite(Vessel_types_n_names, 
       here::here("data", "ship_types_n_names.csv"),
       bom = TRUE)

```


## Vessel filtering

```{r filterVesselbyType}

filter_vessel_type <- function(df, vessel_type){
  
  filtered_df <- df %>% 
    filter(.data[['ShipType']] == vessel_type)
  
  return(filtered_df)
}

tic(msg = "begin filtering")
check_filter_df <- filter_vessel_type(lazy_df_ships, sample_ship_t_n$ShipType[4]) %>%  #Fishing 
  as_tibble()
toc()
```

```{r getShipNames}
get_ship_names <- function(df, vessel_type){
    vessel_names <- df %>% 
    filter(.data[['ShipType']] == vessel_type) %>% 
    select(.data[['ShipName']]) %>% 
    distinct() %>% 
    as_tibble()
    
    return(vessel_names)
}

ship_names <- get_ship_names(Vessel_types_n_names, sample_ship_t_n$ShipType[4])

```

```{r filterbyShipName}
filter_vessel_name <- function(df, vessel_name){
  
  filtered_df <- df %>% 
    filter(.data[['ShipName']] == vessel_name)
  
  return(filtered_df)
}

check_filter_name_df <- filter_vessel_name(lazy_df_ships, "KERLI - 3338") %>% 
  as_tibble()
```


Test filtered df contains same amount of columns

Test expect distinct(ShipType, ShipName) has `r nrow(Vessel_types_n_names)` rows (1,226).

Test some expected answers that were precalculated regarding
distinct() combinations of `ShipType` and `ShipName`

### On selector

### Selecting the longest route given 2 points

```{r caseSelect}

#arrange data by DATETIME then perform rolling/window calculation

case_Kerli <- filter_vessel_name(lazy_df_ships, "KERLI - 3338") %>% 
  arrange(DATETIME) %>% 
  as_tibble()

kerli_datime_n_lat_lon <- case_Kerli %>% select(DATETIME, LAT, LON)
```

### What distance is calculated?

#### Great Circle

[Great circle distances](https://www.r-bloggers.com/2020/02/three-ways-to-calculate-distances-in-r/)

 >>find the observation when it sailed the longest distance between two consecutive observations.

##### Function creation

```{r keyFunctions4Task}
library(sf)
#library(furrr)

# https://r-spatial.github.io/sf/reference/geos_measures.html: great circle distance calculations use by default spherical distances

#transform data to spatial df to perform operations
transform_to_spatial_df <- function(df) {
  if (class(df)[1] != "tbl_df") {
    print("object is not of class tbl_df. please make your df a tibble")
    
  } else{
    spatial_df <- df %>%
      select(DATETIME, .data[['LAT']], .data[['LON']]) %>%
      st_as_sf(.,
               coords = c("LON", "LAT"),
               crs = 4326) %>%
      arrange(DATETIME)
    
    return(spatial_df)
  }
}

#CHECK its a spatial df wth crs 4326

#function distace to use
func_distance <- function(from, to) {
  st_distance(from, to, by_element = TRUE)
}

#check it returns a vector ad that from and to have geometry and crs 4326

#calculate a new tibble after applied function
calc_circle_distance <- function(spatial_df) {
  
  if (class(spatial_df)[1] != "sf") {
    print("not an sf object. Lat and lon features must be encoded as geometries in an sf object.")
    
  } else{
    
    #grab the observation after the current one
    lead_df <- spatial_df %>% lead()
    
    df_distances <- tibble(
      datetime_from = spatial_df$DATETIME,
      datetime_to = lead_df$DATETIME,
      #calculate the distance between poit A from to point B to
      distance_travelled =  func_distance(from = spatial_df, to = lead_df)
    )
    
    return(df_distances)
  }
  
}

#CHECK it receives a spatial df and

get_max_distance <- function(df_distances) {
  if (colnames(df_distances)[3] != "distance_travelled") {
    print("Column needed for operation missing. `distance travelled` is needed")
    
  } else{
    top_distance <- df_distances  %>%
      #HOW TO PROTECT AGAINST SAME DISTANCE AND GRAB THE MOST RECENT?
      slice_max(order_by = distance_travelled)
    
    return(top_distance)
  }
}

obtain_record_of_top_distance <- function(df, top_distance){
  
record_top_distance_points <- tibble(
  
  df %>% 
    filter(DATETIME == top_distance$datetime_from |
             DATETIME == top_distance$datetime_to)
  ) 
  
  return(record_top_distance_points)
}

#CHECK RETURNS ONLY 2 VALUES
```

##### test env

```{r testEnv}
sf_kerli <- sf::st_as_sf(kerli_datime_n_lat_lon, coords = c("LON", "LAT"),
                         crs = 4326) # to get the units in meters

lead_kerli <- sf_kerli %>% arrange(DATETIME) %>% lead()

# set up parraller processing
#plan(multisession, workers = 2) #my windows laptop only has 2 corsd

tic(msg = "sf simple calculation")
kerli_distances <- tibble(
  datetime_from = sf_kerli$DATETIME,
  datetime_to = lead_kerli$DATETIME,
  distance_travelled =  func_distance(from = sf_kerli, to = lead_kerli) # if TRUE, return a vector with distance between the first elements of x and y
) %>% 
  group_by(datetime_from, distance_travelled) %>% 
  slice_max(order_by = distance_travelled)
  #arrange(desc(distance_travelled), DATETIME)
toc()

```

##### full calculation

```{r applyfullcalculation}
tic(msg = "sf integrated calculation")
tic("transform_to_spatial_df")
spatial_kerli <- transform_to_spatial_df(case_Kerli)
toc()
tic("calc_circle_distance")
distances_kerli <- calc_circle_distance(spatial_kerli)
toc()
tic("get max distance travelled")
top_kerli_distance <- get_max_distance(distances_kerli)
toc()
toc()


tic(msg = "sf integrated calculation as a pipe")
tic(msg = "pipe 2 get the answer")
top_distance_kerli <- case_Kerli %>% 
 transform_to_spatial_df(.) %>% 
  calc_circle_distance(.) %>% 
  get_max_distance(.)
toc()
tic(msg = "obtain record that belong to that datetime")
top_record_kerli_distance <- obtain_record_of_top_distance(case_Kerli, top_distance_kerli)
toc()
toc()
```



## Plot results in leaflet

```{r leaftletmap}
library(leaflet)

plot_df <- top_record_kerli_distance %>% 
  arrange(DATETIME) %>% 
  mutate(
    position = c("Point A", "Point B")
  )

leaflet(data = plot_df) %>% addTiles() %>%
  addMarkers(~LON, ~LAT,
             #message to be displayed on click
             popup = ~paste(sep = "<br/>",
                                    ShipName,
                            ShipType,
DATETIME,
position),
             #to display a text label either on hover or statically
             label = ~as.character(position),
labelOptions = labelOptions(noHide = T,
                            textsize = "15px")) %>%
  addMiniMap() %>% 
  addPolylines(~LON, ~LAT,
               label = ~paste(sep = " ", "Distance travelled:", prettyNum(top_distance_kerli$distance_travelled, big.mark = ","),
                              "meters")
               )


```
### full calculation starting from a filtered df

```{r fullmapcalc}
source("R/func_create_map.R")

tic()
create_distance_map_between_a_n_b(case_Kerli)
toc()
```


## Mock answer

The preloaded configuration of the app will be:

  - Vessel type: High Special - 4
  - Vessen name: POLIS 593790-323857
  
```{r coockedCalculation}

```


